\documentclass[12pt]{article}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{mathtools}
\usepackage{hyperref}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{assumption}{Assumption}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\title{Theorem 7: Universal Approximation with Quantitative Rates\\
\large Lean~4 / Mathlib Formal Verification Notes}
\author{Maolin Wang}
\date{\today}

\begin{document}
\maketitle

\section{Assumptions}

\begin{assumption}[Regularity]\label{assumption:regularity}
$F_0(\tau|\boldsymbol{h})$ is a valid CDF in $\tau$ for every $\boldsymbol{h}\in\mathcal{H}$:
monotone, $F_0(0|\boldsymbol{h})=0$, $F_0(\tau|\boldsymbol{h})\to 1$ as $\tau\to\infty$,
and $F_0(\tau|\boldsymbol{h})\in[0,1]$.
\end{assumption}

\begin{assumption}[Architecture]\label{assumption:architecture}
The MNN has non-negative $\tau$-path weights, output via $\sigma(\mathcal{M})$.
\end{assumption}

\begin{assumption}[Smoothness]\label{assumption:smoothness}
$F_0$ is $\alpha$-H\"older in $\boldsymbol{h}$, uniform in $\tau$:
there exist $C,\alpha>0$ such that
\[
  |F_0(\tau|\boldsymbol{h}_1) - F_0(\tau|\boldsymbol{h}_2)|
  \;\le\; C\,\|\boldsymbol{h}_1 - \boldsymbol{h}_2\|^{\alpha}
  \quad \forall\,\tau\ge 0,\;\boldsymbol{h}_1,\boldsymbol{h}_2\in\mathcal{H}.
\]
\end{assumption}

\section{Main Theorem}

\begin{theorem}[Universal Approximation with Quantitative Rates]
\label{thm:universal_approx}
Under Assumptions~\ref{assumption:regularity}--\ref{assumption:architecture}
and~\ref{assumption:smoothness}, for any $\epsilon > 0$, there exists a
network configuration $\theta^*$ with
\[
  N^* = O\!\left(\epsilon^{-(1+d/\alpha)}\right) \text{ parameters}
\]
such that
\[
  \sup_{\tau \geq 0,\, \boldsymbol{h} \in \mathcal{H}}
  \bigl|F^*(\tau|\boldsymbol{h}) - F_0(\tau|\boldsymbol{h})\bigr| < \epsilon,
\]
where $F^*(\tau|\boldsymbol{h}) = \sigma\!\bigl(\mathcal{M}(\tau,\boldsymbol{h};\theta^*)\bigr)$
is a valid CDF for all~$\theta^*$.
\end{theorem}

\begin{proof}
\emph{This is an existence proof}: we construct a specific network in the
hypothesis class satisfying Assumption~\ref{assumption:architecture} that
achieves the stated bound. The proof proceeds in two stages.
\textbf{Stage~A} (Steps~1--6) builds an auxiliary CDF approximation
$\hat{\mathcal{M}} \in [0,1]$ satisfying
$|\hat{\mathcal{M}} - F_0| < \epsilon/2$ everywhere.
\textbf{Stage~B} (Step~7) lifts $\hat{\mathcal{M}}$ into the
$\sigma$-based MNN architecture, incurring at most $\epsilon/2$ additional error.

%% ============ STAGE A ============
\medskip\noindent\textsc{Stage A: Direct CDF Approximation.}

\textbf{Step~1 (Tail and boundary reduction).}
By Lemma~\ref{lem:tail}, choose $T^* = T^*(\epsilon/8,\mathcal{H})$ so that
\[
  \inf_{\boldsymbol{h}\in\mathcal{H}} F_0(T^*|\boldsymbol{h}) > 1 - \tfrac{\epsilon}{8}.
\]
Since $F_0(0|\boldsymbol{h}) = 0$ and $F_0$ is continuous (Assumption~\ref{assumption:regularity}),
choose $\tau_{\min} > 0$ such that
\[
  \sup_{\boldsymbol{h}\in\mathcal{H}} F_0(\tau_{\min}|\boldsymbol{h}) < \tfrac{\epsilon}{8}.
\]
The core region is $[\tau_{\min},\,T^*] \times \mathcal{H}$,
on which $F_0$ takes values in $[c_{\ell},\, 1-c_u]$ for constants
$c_{\ell}, c_u > 0$ independent of~$\epsilon$.

\textbf{Step~2 (History discretisation).}
Define the \emph{Gaussian softmax $\alpha$-moment}
\[
  \Gamma_{d,\alpha}
  \;:=\;
  \sup_{\boldsymbol{h}\in\mathcal{H}}\;
  \sum_{j} w_j(\boldsymbol{h})\,
  \Bigl(\frac{\|\boldsymbol{h}_j - \boldsymbol{h}\|}{\delta}\Bigr)^{\!\alpha},
\]
where $w_j$ are the Gaussian softmax weights defined in Step~4 below.
This constant is finite: the nearest centre $j^*$ contributes $\le 1$,
while a centre at distance $r\delta$ ($r>1$) receives softmax weight
$w_j \le e^{1-r^2}$,
and the number of centres in the shell $[r\delta,(r\!+\!1)\delta]$
is $O(r^{d-1})$.
Hence
$\Gamma_{d,\alpha}
  \le 1 + e\!\sum_{r=1}^{\infty} C' r^{d-1+\alpha} e^{-r^2} < \infty$.

Set
$\delta = \bigl(\epsilon\,/\,(8\,C\,\max\{1,\Gamma_{d,\alpha}\})\bigr)^{1/\alpha}$.
Cover the compact set $\mathcal{H}\subset\mathbb{R}^d$ with a $\delta$-net
$\{\boldsymbol{h}_j\}_{j=1}^K$ where
$K = O(\epsilon^{-d/\alpha})$.

\textbf{Step~3 (Pointwise monotone approximation).}
For each centre $\boldsymbol{h}_j$, apply a piecewise-linear construction to
$g_j(\tau) := F_0(\tau|\boldsymbol{h}_j)$ on $[0,T^*]$ with tolerance $\epsilon/8$.
This yields a non-decreasing function
$\hat{g}_j$ with $O(\epsilon^{-1/\alpha})$ breakpoints satisfying
$\|\hat{g}_j - g_j\|_{\infty,[0,T^*]} < \epsilon/8$.

\textbf{Step~4 (Blending).}
Define Gaussian softmax weights:
\[
  w_j(\boldsymbol{h})
  = \frac{\exp\!\bigl(-\|\boldsymbol{h}-\boldsymbol{h}_j\|^2/\delta^2\bigr)}
         {\sum_k \exp\!\bigl(-\|\boldsymbol{h}-\boldsymbol{h}_k\|^2/\delta^2\bigr)},
  \qquad \sum_j w_j \equiv 1,\; w_j \ge 0,
\]
and set
$\hat{\mathcal{M}}(\tau,\boldsymbol{h})
  := \sum_j w_j(\boldsymbol{h})\,\hat{g}_j(\tau)$.
Key properties: (a)~monotone in $\tau$; (b)~range in $[0,1]$; (c)~$\hat{\mathcal{M}}(0,\cdot)=0$.

\textbf{Step~5 (Core error bound on $[0,T^*]\times\mathcal{H}$).}
\begin{align*}
  |\hat{\mathcal{M}} - F_0|
  &\le \sum_j w_j\,|\hat{g}_j - g_j|
       + \sum_j w_j\,|g_j - F_0| \\
  &\le \frac{\epsilon}{8}
       + C\delta^{\alpha}\,\Gamma_{d,\alpha}
  \;\le\; \frac{\epsilon}{4}.
\end{align*}

\textbf{Step~6 (Boundary and tail regions).}
Upper tail ($\tau > T^*$): $|\hat{\mathcal{M}} - F_0| \le 3\epsilon/8 < \epsilon/2$.
Lower boundary ($\tau \in [0,\tau_{\min}]$): $|\hat{\mathcal{M}} - F_0| < 3\epsilon/8 < \epsilon/2$.

Combined:
$\sup_{\tau\ge 0,\,\boldsymbol{h}\in\mathcal{H}}|\hat{\mathcal{M}}-F_0| < \epsilon/2$.

%% ============ STAGE B ============
\medskip\noindent\textsc{Stage B: Lifting to the $\sigma$-MNN Architecture.}

\textbf{Step~7.}
Let $\eta = c_\ell/2 > 0$ (fixed constant, independent of $\epsilon$).
Define $\tilde{\mathcal{M}} = (1-2\eta)\hat{\mathcal{M}} + \eta \in [\eta,1-\eta]$.
Apply the logit lift $\mathcal{M}_{\mathrm{pre}} = \sigma^{-1}(\tilde{\mathcal{M}})$,
re-discretize using the same construction (Steps~2--4) in the pre-sigmoid space
with H\"older constant $C' = C/(\eta(1-\eta))$,
obtaining $\mathcal{M}_{\mathrm{net}}$ with $O(\epsilon^{-(1+d/\alpha)})$ parameters.
Since $\operatorname{Lip}(\sigma)=1/4$:
\[
  |\sigma(\mathcal{M}_{\mathrm{net}}) - \tilde{\mathcal{M}}| < \epsilon/2.
\]

\emph{Final combination:}
\[
  |F^* - F_0|
  \;\le\;
  \underbrace{|\sigma(\mathcal{M}_{\mathrm{net}}) - \tilde{\mathcal{M}}|}_{<\,\epsilon/2}
  +
  \underbrace{|\tilde{\mathcal{M}} - \hat{\mathcal{M}}|}_{\le\,\eta}
  +
  \underbrace{|\hat{\mathcal{M}} - F_0|}_{<\,\epsilon/2 - \eta}
  \;<\; \epsilon.
\]

\textbf{Parameter count:}
$K \times O(\epsilon^{-1/\alpha}) = O(\epsilon^{-(1+d/\alpha)})$.
\end{proof}

\section{Lean~4 Verification Summary}

The file \texttt{Thm7.lean} formalizes each step as a separate lemma in Lean~4/Mathlib~4.

\subsection*{Fully Proved Lemmas (no \texttt{sorry})}

\begin{center}
\begin{tabular}{ll}
\hline
\textbf{Lemma} & \textbf{Key tactic} \\
\hline
\texttt{step1\_tail\_bdy} & \texttt{max} construction + monotonicity \\
\texttt{gaussW\_nonneg} & \texttt{div\_nonneg} + \texttt{exp\_pos} \\
\texttt{gaussW\_sum\_one} & \texttt{Finset.sum\_div} + \texttt{div\_self} \\
\texttt{step4a\_mono} & \texttt{Finset.sum\_le\_sum} + convexity \\
\texttt{step4b\_range} & Convex combination of $[0,1]$-valued functions \\
\texttt{step4c\_boundary} & \texttt{Finset.sum\_eq\_zero} + \texttt{mul\_zero} \\
\texttt{step6\_upper\_tail} & \texttt{abs\_lt} + \texttt{linarith} \\
\texttt{step6\_lower\_bdy} & \texttt{abs\_lt} + \texttt{linarith} \\
\texttt{perturbation\_range} & \texttt{nlinarith} \\
\texttt{step7\_combine} & \texttt{rw [abs\_lt]} + \texttt{linarith} (ring identity) \\
\hline
\end{tabular}
\end{center}

\subsection*{Lemmas Using \texttt{sorry} (Deep Analysis)}

\begin{itemize}
  \item \texttt{step2\_Gamma\_finite} — Gaussian shell summation: $\Gamma_{d,\alpha} < \infty$.
  \item \texttt{step3\_pwl} — PWL approximation lemma.
    \emph{Note: requires $\tau$-direction H\"older regularity of $F_0$, which is
    not stated in the current assumptions and must be added.}
  \item \texttt{step5\_core\_error} — Weighted triangle inequality over softmax mixture.
  \item \texttt{sigmoid\_quarter\_lip} — $\operatorname{Lip}(\sigma)=1/4$; proof uses
    \texttt{lipschitzWith\_of\_nnnorm\_deriv\_le} but NNReal coercion API differs
    across Mathlib versions.
  \item \texttt{logit\_lip\_on\_interval} — MVT on $\sigma^{-1}$.
  \item \texttt{stage\_A}, \texttt{stage\_B}, \texttt{theorem7} — Orchestration skeleton.
\end{itemize}

\subsection*{Known Issues / Paper Gaps}

\begin{enumerate}
  \item \textbf{Step~2 circular dependency:} $\Gamma_{d,\alpha}$ is defined via $w_j$
    which depends on $\delta$, but $\delta$ depends on $\Gamma_{d,\alpha}$.
    Resolved by noting the normalized distances $\|\boldsymbol{h}_j-\boldsymbol{h}\|/\delta$
    form a unit-spaced lattice independent of $\delta$, but this argument
    should be made explicit.

  \item \textbf{Step~3 missing assumption:} The PWL approximation requires
    $\tau$-direction H\"older regularity of $F_0$.
    Currently only $\boldsymbol{h}$-direction smoothness is assumed.

  \item \textbf{Step~7 tolerance adjustment:} Stage~A must be run with tolerance
    $\epsilon/2 - \eta$ (not $\epsilon/2$) for the final error to sum to $\epsilon$.
    The proof mentions this but could be made more explicit.
\end{enumerate}

\end{document}
